Resurgence of the Airy function

### 2021-12-14 airy-resurgence.tex
Transform equation using translation and scaling operators

%%Say $A$ is an $n$th-order differential operator with polynomial coefficients. If we know that $A\hat{\varphi} = 0$, and we know the Taylor polynomial of $\hat{\varphi}$ at zero through degree $n - 1$, we can use the formulas above to find an inhomogeneous differential equation that $\laplace \hat{\varphi}$ must satisfy.
\subsection{Algebraic version}
\subsubsection{Definition}
Let $\mathcal{P}$ be the vector space spanned by $\zeta^\alpha$ for $\alpha \in \R \smallsetminus \Z_{< 0}$. Note that $\mathcal{P} \cap \mathcal{E}$ is $\mathcal{P}_{> -1}$, the subspace spanned by $\zeta^\alpha$ with $\alpha > -1$. Since $\laplace(\zeta^\alpha) = \Gamma(\alpha+1)\,z^{-(\alpha + 1)}$ for all $\alpha > -1$, let's use the same formula to extend $\laplace$ to all of $\mathcal{P}$. This defines $\laplace$ consistently on $\mathcal{E} + \mathcal{P}$.
\subsubsection{Action on differential operators}
Observe that
\[ \laplace(\zeta^{\alpha + 1}) = -\tfrac{\partial}{\partial z}\,\laplace(\zeta^\alpha) \]
for $\alpha \neq -1$. This extends identity~\ref{id:L-mult} to all of $\mathcal{P}$.

@@

Now we have the results of Section~\ref{L-diff-op} for all functions in $\mathcal{E} + \mathcal{P}$. Identity~\ref{id:L-diff} is particularly simple for functions that can be written as $\hat{\varphi}_\text{frac} + \hat{\varphi}_\text{reg}$, where $\hat{\varphi}_\text{frac} \in \mathcal{P}$ has only non-integer exponents, and the zeroth through $(n-1)$st derivatives of $\hat{\varphi}_\text{reg} \in \mathcal{E}$ vanish at zero. For these functions, all the initial values vanish, and we have
\[ \laplace \big(\tfrac{\partial}{\partial \zeta}\big)^n (\hat{\varphi}_\text{frac} + \hat{\varphi}_\text{reg}) = z^n \laplace (\hat{\varphi}_\text{frac} + \hat{\varphi}_\text{reg}). \]
\section{The Airy equation}
\subsection{Basics}
The Airy equation is
\begin{equation}\label{eqn:airy}
\left[\big(\tfrac{\partial}{\partial x}\big)^2 - x\right] \psi = 0.
\end{equation}
One solution is given by the Airy function,
\[ \Ai(x) = \frac{1}{2\pi i} \int_{\Gamma_0} \exp\left(\tfrac{1}{3}t^3 - xt\right)\,dt, \]
where $\Gamma_0$ is a path that comes from $\infty$ at $-60^\circ$ and goes to $\infty$ at $60^\circ$. With the substitution $t = 2ux^{1/2}$, we can rewrite the Airy integral as
\[ \Ai(x) = x^{1/2}\;\frac{1}{\pi i} \int_{x^{-1/2} \Gamma_0} \exp\left[\tfrac{2}{3}x^{3/2} \left(4u^3 - 3u\right)\right]\,du. \]
We've rescaled the contour by a factor of two, but it still approaches $\infty$ in the desired way. Note that $4u^3 - 3u$ is the third Chebyshev polynomial.
\subsection{Rewriting as a modified Bessel equation}
We can distill the most interesting part of the Airy function by writing
\[ \Ai(x) = \tfrac{1}{\pi\sqrt{3}}\,x^{1/2}\, K\big(\tfrac{2}{3} x^{3/2}\big), \]
where
\[ K(z) = \frac{\sqrt{3}}{i} \int_{z^{-1/3}\Gamma_0} \exp\left[z \left(4u^3 - 3u\right)\right]\,du. \]
Saying that $\Ai$ satisfies the Airy equation is equivalent to saying that $K$ satisfies the modified Bessel equation
\begin{equation}\label{eqn:mod-bessel}
\left[z^2 \big(\tfrac{\partial}{\partial z}\big)^2 + z \tfrac{\partial}{\partial z} - \big[\big(\tfrac{1}{3}\big)^2 + z^2\big]\right] \varphi = 0.
\end{equation}
In fact, $K$ is the modified Bessel function $K_{1/3}$~\cite[equation~9.6.1]{dlmf}.
\subsection{Going to the spatial domain}
Let's find a function $\hat{K}$ with $K = \laplace \hat{K}$, which is unique if it exists~\cite[Theorem~1.23]{laplace-tfm}. If a function $\hat{\varphi}$ satisfies the equation
\begin{equation}\label{eqn:spatial-mod-bessel}
\left[\big(\zeta^2 - 1\big) \big(\tfrac{\partial}{\partial \zeta}\big)^2 + 3\zeta \tfrac{\partial}{\partial \zeta} + \big[1 - \big(\tfrac{1}{3}\big)^2\big]\right] \hat{\varphi} = 0,
\end{equation}
its Laplace transform $\varphi = \laplace \hat{\varphi}$ satisfies the equation
\begin{align*}
\left[\big({-\tfrac{\partial}{\partial z}}\big)^2 - 1\right] \Big[z^2 \varphi - \Big(\hat{\varphi}(0)\,z + \hat{\varphi}'(0)\Big)\Big] + 3\big({-\tfrac{\partial}{\partial z}}\big)\Big[z\varphi - \hat{\varphi}(0)\Big] + \big[1 - \big(\tfrac{1}{3}\big)^2\big] \varphi & = 0 \\
\big(\tfrac{\partial}{\partial z}\big)^2 \big[z^2 \varphi\big] - \Big[z^2 \varphi - \Big(\hat{\varphi}(0)\,z + \hat{\varphi}'(0)\Big)\Big] - 3\big(\tfrac{\partial}{\partial z}\big)\big[z\varphi\big] + \big[1 - \big(\tfrac{1}{3}\big)^2\big] \varphi & = 0 \\
\Big[2 + 4z\tfrac{\partial}{\partial z} + z^2\big(\tfrac{\partial}{\partial z}\big)^2\Big]\varphi - \Big[z^2 \varphi - \Big(\hat{\varphi}(0)\,z + \hat{\varphi}'(0)\Big)\Big] - 3\Big[1 + z\tfrac{\partial}{\partial z}\Big]\varphi + \big[1 - \big(\tfrac{1}{3}\big)^2\big] \varphi & = 0,
\end{align*}
which simplifies to
\begin{equation}\label{eqn:inhomog-mod-bessel}
\left[z^2 \big(\tfrac{\partial}{\partial z}\big)^2 + z \tfrac{\partial}{\partial z} - \big[\big(\tfrac{1}{3}\big)^2 + z^2\big]\right] \varphi = -\Big(\hat{\varphi}(0)\,z + \hat{\varphi}'(0)\Big).
\end{equation}
Noticing that equation~\ref{eqn:inhomog-mod-bessel} is an inhomogeneous version of equation~\ref{eqn:mod-bessel}, we

%%This suggests that $\hat{K}$ could be a solution of equation~\ref{eqn:spatial-mod-bessel} that vanishes through first degree at $\zeta = 0$.

With the change of variable $\xi = \tfrac{1}{2}(1-\zeta)$, equation~\ref{eqn:spatial-mod-bessel} becomes the hypergeometric equation
\begin{equation}\label{eqn:hypergeom}
\left[\xi(1 - \xi) \big(\tfrac{\partial}{\partial \xi}\big)^2 + 3(\tfrac{1}{2} - \xi) \tfrac{\partial}{\partial \xi} - \big[1 - \big(\tfrac{1}{3}\big)^2\big]\right] \hat{\varphi} = 0.
\end{equation}
\textbf{[...]}

The hypergeometric function
\[ F\big(\tfrac{2}{3}, \tfrac{4}{3}; \tfrac{3}{2}; \xi\big) \]
satisfies equation~\ref{eqn:hypergeom} by definition. It has a singularity of type $\xi^{-1/2}$ at $\xi = 0$, and continues holomorphically to the rest of $\C$. Using DLMF~15.10.13, we can see that
\[ F\big(\tfrac{2}{3}, \tfrac{4}{3}; \tfrac{3}{2}; 1-\xi\big). \]
also satisies equation~\ref{eqn:hypergeom}. It has a singularity of type $(1 - \xi)^{-1/2}$ at $\xi = 1$, and continues holomorphically to the rest of $\C$.

>>>>

%%Say $A$ is an $n$th-order differential operator with polynomial coefficients. If we know that $A\hat{\varphi} = 0$, and we know the Taylor polynomial of $\hat{\varphi}$ at zero through degree $n - 1$, we can use the formulas above to find an inhomogeneous differential equation that $\laplace \hat{\varphi}$ must satisfy.
\subsection{Algebraic version}
\subsubsection{Definition}
Let $\mathcal{P}$ be the vector space spanned by $\zeta^\alpha$ for $\alpha \in \R \smallsetminus \Z_{< 0}$. Note that $\mathcal{P} \cap \mathcal{E}$ is $\mathcal{P}_{> -1}$, the subspace spanned by $\zeta^\alpha$ with $\alpha > -1$. Since $\laplace(\zeta^\alpha) = \Gamma(\alpha+1)\,z^{-(\alpha + 1)}$ for all $\alpha > -1$, let's use the same formula to extend $\laplace$ to all of $\mathcal{P}$. This defines $\laplace$ consistently on $\mathcal{E} + \mathcal{P}$.
\subsubsection{Action on differential operators}\label{L-diff-op-alg}
Observe that
\[ \laplace(\zeta^{\alpha + 1}) = -\tfrac{\partial}{\partial z}\,\laplace(\zeta^\alpha) \]
for $\alpha \neq -1$. This extends identity~\ref{id:L-mult} to all of $\mathcal{P}$.

@@

Now we have the results of Section~\ref{L-diff-op} for all functions in $\mathcal{E} + \mathcal{P}$. Identity~\ref{id:L-diff} is particularly simple for functions that can be written as $\hat{\varphi}_\text{frac} + \hat{\varphi}_\text{reg}$, where $\hat{\varphi}_\text{frac} \in \mathcal{P}$ has only non-integer exponents, and the zeroth through $(n-1)$st derivatives of $\hat{\varphi}_\text{reg} \in \mathcal{E}$ vanish at zero. For these functions, all the initial values vanish, and we have
\[ \laplace \big(\tfrac{\partial}{\partial \zeta}\big)^n (\hat{\varphi}_\text{frac} + \hat{\varphi}_\text{reg}) = z^n \laplace (\hat{\varphi}_\text{frac} + \hat{\varphi}_\text{reg}). \]
\subsection{Interactions with other operators}
\subsubsection{Translation}
Let $T_a$ be the left translation operator $[T_a f](\zeta) = f(\zeta + a)$. The Laplace transform commutes between $T_a$ and multiplication by $e^{az}$:
\[ \laplace T_a \hat{\varphi} = e^{az} \laplace \hat{\varphi}. \]
Consistent with Section~\ref{L-diff-op}, we have
\begin{align*}
(\zeta + a) T_a \hat{\varphi} & = T_a(\zeta \hat{\varphi}) \\
\big(\tfrac{\partial}{\partial z} - a\big) e^{az} \varphi & = e^{az} \tfrac{\partial}{\partial z} \varphi
\end{align*}
and
\begin{align*}
\tfrac{\partial}{\partial \zeta} T_a & = T_a \tfrac{\partial}{\partial \zeta} \\
z e^{az} \varphi & = e^{az} z \varphi.
\end{align*}
\subsubsection{Scaling}
Let $S^k_c$ be the homogeneous scaling operator $[S^k_c f](\zeta) = c^k f(c \zeta)$. The Laplace transform commutes between $S^1_c$ and $S^0_{c^{-1}}$:
\[ \laplace S^1_c \hat{\varphi} = S^0_{c^{-1}} \laplace \hat{\varphi}. \]
Consistent with Section~\ref{L-diff-op}, we have
\begin{align*}
c\zeta S^k_c \hat{\varphi} & = S^k_c(\zeta \hat{\varphi}) \\
c \tfrac{\partial}{\partial z} S^k_{c^{-1}} \varphi & = S^k_{c^{-1}} \tfrac{\partial}{\partial z} \varphi.
\end{align*}
\section{The Airy equation}
\subsection{Basics}
The Airy equation is
\begin{equation}\label{eqn:airy}
\left[\big(\tfrac{\partial}{\partial x}\big)^2 - x\right] \psi = 0.
\end{equation}
One solution is given by the Airy function,
\[ \Ai(x) = \frac{1}{2\pi i} \int_{\Gamma_0} \exp\left(\tfrac{1}{3}t^3 - xt\right)\,dt, \]
where $\Gamma_0$ is a path that comes from $\infty$ at $-60^\circ$ and goes to $\infty$ at $60^\circ$. With the substitution $t = 2ux^{1/2}$, we can rewrite the Airy integral as
\[ \Ai(x) = x^{1/2}\;\frac{1}{\pi i} \int_{x^{-1/2} \Gamma_0} \exp\left[\tfrac{2}{3}x^{3/2} \left(4u^3 - 3u\right)\right]\,du. \]
We've rescaled the contour by a factor of two, but it still approaches $\infty$ in the desired way. Note that $4u^3 - 3u$ is the third Chebyshev polynomial.
\subsection{Rewriting as a modified Bessel equation}
We can distill the most interesting part of the Airy function by writing
\[ \Ai(x) = \tfrac{1}{\pi\sqrt{3}}\,x^{1/2}\, K\big(\tfrac{2}{3} x^{3/2}\big), \]
where
\[ K(z) = \frac{\sqrt{3}}{i} \int_{z^{-1/3}\Gamma_0} \exp\left[z \left(4u^3 - 3u\right)\right]\,du. \]
Saying that $\Ai$ satisfies the Airy equation is equivalent to saying that $K$ satisfies the modified Bessel equation
\begin{equation}\label{eqn:mod-bessel}
\left[z^2 \big(\tfrac{\partial}{\partial z}\big)^2 + z \tfrac{\partial}{\partial z} - \big[\big(\tfrac{1}{3}\big)^2 + z^2\big]\right] \varphi = 0.
\end{equation}
In fact, $K$ is the modified Bessel function $K_{1/3}$~\cite[equation~9.6.1]{dlmf}.

As we'll see in Section~\ref{asymp}, $K$ is in $O_\infty(e^{-z})$. It'll be helpful to pull out the exponential decay factor and work instead with the function $\kappa$ defined by $K(z) = e^{-z} \kappa(z)$. Saying that $K$ satisfies equation~\ref{eqn:mod-bessel} is equivalent to saying that $\kappa$ satisfies the equation
\begin{equation}\label{eqn:shifted-mod-bessel}
\left[z^2 \big(\tfrac{\partial}{\partial z} + 1\big)^2 + z \big(\tfrac{\partial}{\partial z} + 1\big) - \big[\big(\tfrac{1}{3}\big)^2 + z^2\big]\right] \varphi = 0.
\end{equation}
%%\subsection{Asymptotic analysis}\label{asymp}
\subsection{Going to the spatial domain}
\subsubsection{First try}
Let's try to find a function $\hat{K}$ with $K = \laplace \hat{K}$, which is unique if it exists~\cite[Theorem~1.23]{laplace-tfm}. If a function $\hat{\varphi}$ satisfies the equation
\begin{equation}\label{eqn:spatial-mod-bessel}
\left[\big(\zeta^2 - 1\big) \big(\tfrac{\partial}{\partial \zeta}\big)^2 + 3\zeta \tfrac{\partial}{\partial \zeta} + \big[1 - \big(\tfrac{1}{3}\big)^2\big]\right] \hat{\varphi} = 0,
\end{equation}
its Laplace transform $\varphi = \laplace \hat{\varphi}$ satisfies the equation
\begin{align*}
\left[\big({-\tfrac{\partial}{\partial z}}\big)^2 - 1\right] \Big[z^2 \varphi - \Big(\hat{\varphi}(0)\,z + \hat{\varphi}'(0)\Big)\Big] + 3\big({-\tfrac{\partial}{\partial z}}\big)\Big[z\varphi - \hat{\varphi}(0)\Big] + \big[1 - \big(\tfrac{1}{3}\big)^2\big] \varphi & = 0 \\
\big(\tfrac{\partial}{\partial z}\big)^2 \big[z^2 \varphi\big] - \Big[z^2 \varphi - \Big(\hat{\varphi}(0)\,z + \hat{\varphi}'(0)\Big)\Big] - 3\big(\tfrac{\partial}{\partial z}\big)\big[z\varphi\big] + \big[1 - \big(\tfrac{1}{3}\big)^2\big] \varphi & = 0 \\
\Big[2 + 4z\tfrac{\partial}{\partial z} + z^2\big(\tfrac{\partial}{\partial z}\big)^2\Big]\varphi - \Big[z^2 \varphi - \Big(\hat{\varphi}(0)\,z + \hat{\varphi}'(0)\Big)\Big] - 3\Big[1 + z\tfrac{\partial}{\partial z}\Big]\varphi + \big[1 - \big(\tfrac{1}{3}\big)^2\big] \varphi & = 0,
\end{align*}
which simplifies to
\begin{equation}\label{eqn:inhomog-mod-bessel}
\left[z^2 \big(\tfrac{\partial}{\partial z}\big)^2 + z \tfrac{\partial}{\partial z} - \big[\big(\tfrac{1}{3}\big)^2 + z^2\big]\right] \varphi = -\Big(\hat{\varphi}(0)\,z + \hat{\varphi}'(0)\Big).
\end{equation}
Since we want $\laplace \hat{K}$ to satisfy equation~\ref{eqn:mod-bessel}, which is the homogeneous version of equation~\ref{eqn:inhomog-mod-bessel}, we might guess that $\hat{K}$ is a solution of equation~\ref{eqn:spatial-mod-bessel} that vanishes through first order at zero. Unfortunately, this would force $\hat{K}$ to be zero.
\subsubsection{Success}
Let's instead try to find a function $\hat{\kappa}$ with $\laplace \hat{\kappa} = \kappa$, or equivalently $\hat{\kappa} = T_1 \hat{K}$. Applying $T_1$ to both sides of equation~\ref{eqn:spatial-mod-bessel}, we get the equivalent equation
\begin{equation}\label{eqn:shifted-spatial-mod-bessel}
\left[\zeta(\zeta + 2) \big(\tfrac{\partial}{\partial \zeta}\big)^2 + 3(\zeta + 1) \tfrac{\partial}{\partial \zeta} + \big[1 - \big(\tfrac{1}{3}\big)^2\big]\right] T_1 \hat{\varphi} = 0.
\end{equation}
If $T_1 \hat{\varphi}$ satisfies equation~\ref{eqn:shifted-spatial-mod-bessel}, $e^z \laplace \hat{\varphi}$ will satisfy an inhomogeneous version of equation~\ref{eqn:shifted-mod-bessel}, analogous to equation~\ref{eqn:inhomog-mod-bessel}. This time, though, there's a trick we can use to zero out the inhomogeneity. Equation~\ref{eqn:shifted-spatial-mod-bessel} has a regular singularity at zero, with one solution line generated by a holomorphic multiple of $\zeta^{-1/2}$. The solutions on that line can be written as $\hat{\varphi}_\text{frac} + \hat{\varphi}_\text{reg}$ as described Section~\ref{L-diff-op-alg}, so their Laplace transforms satisfy equation~\ref{eqn:shifted-mod-bessel}.

Following this plan, let's find $\hat{\kappa}$ explicitly. Applying $S^1_{-2}$ to both sides of equation~\ref{eqn:shifted-spatial-mod-bessel}, we get the equivalent equation
\begin{align*}
\left[(-2\zeta)(-2\zeta + 2) \big({-\tfrac{1}{2}} \tfrac{\partial}{\partial \zeta}\big)^2 + 3(-2\zeta + 1) \big({-\tfrac{1}{2}}\tfrac{\partial}{\partial \zeta}\big) + \big[1 - \big(\tfrac{1}{3}\big)^2\big]\right] S^1_{-2} T_1 \hat{\varphi} & = 0 \\
\left[\zeta(\zeta - 1) \big(\tfrac{\partial}{\partial \zeta}\big)^2 + 3(\zeta - \tfrac{1}{2}) \tfrac{\partial}{\partial \zeta} + \big[1 - \big(\tfrac{1}{3}\big)^2\big]\right] S^1_{-2} T_1 \hat{\varphi} & = 0,
\end{align*}
which simplifies to the hypergeometric equation
\begin{equation}\label{eqn:hypergeom}
\left[\zeta(1 - \zeta) \big(\tfrac{\partial}{\partial \zeta}\big)^2 + 3(\tfrac{1}{2} - \zeta) \tfrac{\partial}{\partial \zeta} - \big[1 - \big(\tfrac{1}{3}\big)^2\big]\right] S^1_{-2} T_1 \hat{\varphi} = 0.
\end{equation}
%With the change of variable $\xi = \tfrac{1}{2}(1-\zeta)$, equation~\ref{eqn:spatial-mod-bessel} becomes the hypergeometric equation
%%\begin{equation}\label{eqn:hypergeom}
%%\left[\xi(1 - \xi) \big(\tfrac{\partial}{\partial \xi}\big)^2 + 3(\tfrac{1}{2} - \xi) \tfrac{\partial}{\partial \xi} - \big[1 - \big(\tfrac{1}{3}\big)^2\big]\right] \hat{\varphi} = 0.
%%\end{equation}
The hypergeometric function
\[ F\big(\tfrac{2}{3}, \tfrac{4}{3}; \tfrac{3}{2}; \xi\big) \]
satisfies equation~\ref{eqn:hypergeom} by definition. It has a singularity of type $\xi^{-1/2}$ at $\xi = 0$, and continues holomorphically to the rest of $\C$. Using DLMF~15.10.13, we can see that
\[ F\big(\tfrac{2}{3}, \tfrac{4}{3}; \tfrac{3}{2}; 1-\xi\big). \]
also satisies equation~\ref{eqn:hypergeom}. It has a singularity of type $(1 - \xi)^{-1/2}$ at $\xi = 1$, and continues holomorphically to the rest of $\C$.
